Probabilistic Symbol Encoding for \\Convolutional Associative Memories

### 2017-01-19 papers/hrr-paper/paper.tex
Add a few changes from reviewer

	% As a general rule, do not put math, special symbols or citations
	% in the abstract
	\begin{abstract}
		Vector Symbolic Architectures (VSAs) define a set of operations
		for association, (compressed) storage, manipulation and retrieval
		of symbolic concepts, represented as fixed-length vectors in
		${\rm I\!R}^n$.
		A specific instance of VSAs, Holographic Reduced Representations (HRRs),
		have proven to exhibit properties similar to human short-term
		memory and are, as such, interesting for computational modeling.
		In this paper we extend the HRR approach by introducing implicit,
		topology-preserving encoding and decoding procedures.
		We propose moving away from traditional exact symbolic representation
		to probabilistic one, replacing unique symbol identifiers with 
		probability density function.
		As we discover, this requires randomly chosen permutations for encoding and
		their corresponding inverse for decoding. These are necessary
		for the uniform distribution of signals across Fourier space where
		embedding takes place.
		These novel encoding schemes, together with the extended operation
		pipeline, are eliminating the need for so-called \emph{clean-up modules} after memory
		retrieval (e.g., self-organizing maps). Effectively, encoding
		implicitly represents the scalar symbol, so no further lookup is needed.
		We show that our encoding scheme has a positive impact
		on memory capacity in comparison to the original capacity
		benchmark for HRRs by Plate
		\cite{Plate:1995:HolographicReducedRepresentations}.
		We also evaluate our memories in two different robotics tasks: visual scene memory and state machine scripting (holographic controllers).
	\end{abstract}

@@

	In the case of lookup tables, in order to decode a specific symbolic vector, the elements from a symbol table are iterated over, the similarity to each of the entries is calculated and the symbol which presented maximum similarity is returned. With the help of a threshold for the similarity value false-positives can be excluded. A lookup requires $O(n)$, where $n$ is the number of symbols known to the system. This approach works quite well for object identifiers (names), but has some non-desirable properties when used on scalar encoding. The space complexity will increase and may become infeasible, depending on the number of inputs needed. One way of handling symbolic scalar encoding is to discretize the whole input range into bins and assigning a random vector to each discrete value. This however may often prove to not satisfy the task needs any more.

@@

The resulting vector of the superposition operation $\pmb m$ can then be probed by using the inverse of circular convolution, which is circular correlation. Plainly put probing can be viewed as binding with the "inverse" of a vector in Fourier space:

>>>>

	% As a general rule, do not put math, special symbols or citations
	% in the abstract
	\begin{abstract}
		Vector Symbolic Architectures (VSAs) define a set of operations
		for association, storage, manipulation and retrieval
		of symbolic concepts, represented as fixed-length vectors in
		${\rm I\!R}^n$.
		A specific instance of VSAs, Holographic Reduced Representations (HRRs),
		have proven to exhibit properties similar to human short-term
		memory and are, as such, interesting for computational modeling.
		In this paper we extend the HRR approach by introducing implicit,
		topology-preserving encoding and decoding procedures.
		We propose moving away from traditional exact symbolic representation
		to probabilistic one, replacing unique symbol identifiers with 
		probability density function.
		As we discover, this requires randomly chosen permutations for encoding and
		their corresponding inverse for decoding. These are necessary
		for the uniform distribution of signals across Fourier space where
		embedding takes place.
		These novel encoding schemes, together with the extended operation
		pipeline, are eliminating the need for so-called \emph{clean-up modules} after memory
		retrieval (e.g., self-organizing maps). Effectively, encoding
		implicitly represents the scalar symbol, so no further lookup is needed.
		We show that our encoding scheme has a positive impact
		on memory capacity in comparison to the original capacity
		benchmark for HRRs \cite{Plate:1995:HolographicReducedRepresentations}.
		We also evaluate our memories in two different robotics tasks: visual scene memory and state machine scripting (holographic controllers).
	\end{abstract}

@@

	In the case of lookup tables, in order to decode a specific symbolic vector, the elements from a symbol table are iterated over, the similarity to each of the entries is calculated and the symbol which presented maximum similarity is returned. With the help of a threshold for the similarity value false-positives can be excluded. A lookup requires $O(n)$, where $n$ is the number of symbols known to the system. This approach works quite well for object identifiers (names), but has some non-desirable properties when used on scalar encoding. The space complexity will increase and may become infeasible, depending on the number of inputs needed. One way of handling symbolic scalar encoding is to discretize the whole input range into bins and assigning a random vector to each discrete value. This however may often prove to not satisfy the task needs any more.
In this paper, we present a probabilistic method for removing the necessity of
using a clean-up procedure after decoding scalar values.

@@

The resulting vector of the superposition operation $\pmb m$ can then be probed
by using the inverse of circular convolution, which is circular correlation.
Probing can be viewed as binding with the "inverse" of a vector in Fourier space:
